---
title: "p8105_hw2_ah3108"
author: "Ava Hamilton"
date: "9/29/2019"
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(kableExtra)
library(arsenal)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%",
  options(scipen = 999)
)

```

# Problem 1

### Read in and clean Mr. Trash data
```{r}

water_df = read_excel(path = "./data_hw2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))

head(water_df)

```

### Read in and clean 2017/2018 precipitation data

```{r}
prec18 = read_excel(path = "./data_hw2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  view

prec17 = read_excel(path = "./data_hw2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2017) %>% 
  view

prec_df = bind_rows(prec17, prec18) %>% 
  mutate(month = month.name[month]) %>% 
  rename(inches_rain = total)

head(prec_df)

```

There were a total of `r nrow(water_df)` observations in the _Mr. Trash Wheel_ dataset, and a total of `r nrow(prec_df)` obersvations when combining the _Precipitation Data_ for 2017 and 2018.

The collection of _Mr. Trash Wheel_ data spans from `r min(pull(water_df, year))` to `r max(pull(water_df, year))`. The average weight of a dumped dumbster was `r round(mean(pull(water_df, weight_tons)), 2)` tons and ranged from `r min(pull(water_df, weight_tons))` to `r max(pull(water_df, weight_tons))` tons. In that time, `r sum(pull(water_df, weight_tons))` tons of trash has powered `r round(sum(pull(water_df, homes_powered)), 2)` homes. Although I personally don't find "sports_balls" to be a **key** variable, the median number of sports balls found per dumpster dump in 2017 was `r median(pull(water_df, sports_balls)[pull(water_df,year) == 2017])`.

_Precipitation data_ showed an average of `r round(mean(pull(prec_df, inches_rain)[pull(prec_df,year) == 2017]), 2)` inches of rain per month in 2017, with a total of `r round(sum(pull(prec_df, inches_rain)[pull(prec_df,year) == 2017]), 2)` inches, and `r round(mean(pull(prec_df, inches_rain)[pull(prec_df,year) == 2018]), 2)` per month in 2018 with a total of `r round(sum(pull(prec_df, inches_rain)[pull(prec_df,year) == 2018]), 2)` inches.


# Problem 2

### Read in and clean FiveThirtyEight pols-month data
```{r message = FALSE}

pols_df = read_csv(file = "./data_hw2/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(col = mon, into = c("year", "month", "day")) %>%
  mutate(year = as.integer(year),
         month = as.integer(month),
         month = month.name[month],
         president = as.character(prez_dem),
         president = recode(president, "1" = "dem", "0" = "gop")
  ) %>% 
  select(-c("prez_dem", "prez_gop", "day")) 
head(pols_df)

```


### Read in and clean FiveThirtyEight snp data
```{r message = FALSE}

snp_df = read_csv(file = "./data_hw2/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(col = date, int = c("month" ,"day", "year")) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         month = month.name[month]) %>% 
  rename(snp_close = close) %>% 
  arrange(year, month) %>% 
  select(year, month, snp_close)
head(snp_df)

```


### Read in and clean FiveThirtyEight unemployment data
```{r message = FALSE}

unemploy_df = read_csv(file = "./data_hw2/fivethirtyeight_datasets/unemployment.csv")

unemploy_long = pivot_longer(
  unemploy_df,
  Jan:Dec,
  names_to = "month",
  values_to = "unemployment")

unemploy_long =   mutate(
  unemploy_long,
  month = match(month, month.abb),
  month = month.name[month]) %>%
  drop_na(unemployment) %>%  
  rename(year = Year) %>%
  arrange(year, month)

head(unemploy_long)

```

### Joining FiveThirtyEight data

```{r}

pols_snp = left_join(pols_df, snp_df, by = c("year", "month"))

all_538_df = left_join(pols_snp, unemploy_long, by = c("year", "month"))
head(all_538_df)

```


The _pols-month_ data gives the number of senators and congress representatives in the Democratic and Republican parties, as well as the party of the president each month from  `r first(pull(pols_df, month), order_by = c(pull(pols_df, year), pull(pols_df, month)))`, `r min(pull(pols_df,year))` to `r last(pull(pols_df, month), order_by = c(pull(pols_df), pull(pols_df, month)))`, `r max(pull(pols_df,year))`.

The _snp_ data gives the closing Standard & Poors market index per date of observation (N = `r nrow(snp_df)`). After editing the data we have an SNP closing market index for the month and year from `r first(pull(snp_df, month), order_by = c(pull(snp_df, year), pull(snp_df, month)))`, `r min(pull(snp_df,year))` to `r last(pull(snp_df, month), order_by = c(pull(snp_df, year), pull(snp_df, month)))`, `r max(pull(snp_df,year))`.

The _unemployment_ data gives the monthly percent of unemployment in the United States from `r first(pull(unemploy_long, month), order_by = c(pull(unemploy_long, year), pull(unemploy_long, month)))`, `r min(pull(unemploy_long,year))` to `r last(pull(unemploy_long, month), order_by = c(pull(unemploy_long), pull(unemploy_long, month)))`, `r max(pull(unemploy_long,year))`.

Given the order in which they are combined we lose data from xx and have NA data in YY.

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).


# Problem 3

### Read in and clean baby names data

```{r message = FALSE}

bn_df = read_csv("./data_hw2/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = as.factor(gender),
    childs_first_name = str_to_upper(childs_first_name),
    childs_first_name = as.factor(childs_first_name),
    ethnicity = as.factor(ethnicity),
    ethnicity = fct_collapse(
      ethnicity,
      ASIAN_PI = c("ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
      BLACK_NH = c("BLACK NON HISP", "BLACK NON HISPANIC"),
      WHITE_NH = c("WHITE NON HISP", "WHITE NON HISPANIC"),
      HISPANIC = "HISPANIC")
  ) %>% 
  rename(year = year_of_birth) %>% 
  
  distinct() %>% 
  view


```
### Creating Table 3.1: Rank of "Olivia" as a First Name, 2011-2016

```{r}

olivia_rank = filter(bn_df, childs_first_name == "OLIVIA") %>%
  select(-c(gender, count, childs_first_name)) %>% 
  arrange(year)

olivia_table  = pivot_wider(olivia_rank,
                            names_from = year,
                            values_from = rank)

kable(olivia_table,  caption = "Rank of 'Olivia', by Year and Race/Ethnicity", col.names = c("", 2011:2016)) %>%  
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) 


```

### Creating Table 3.2: Top First Boy Name 2011-2016

```{r}
boy_one = filter(bn_df, rank == 1 & gender == "MALE") %>% 
  select(-c(gender, count, rank)) %>% 
  arrange(year)

boy_table = pivot_wider(boy_one,
                        names_from = year,
                        values_from = childs_first_name)


kable(boy_table,  caption = "Top Ranked Boy Name from 2011-2016, by Race/Ethnicity", col.names = c("", 2011:2016)) %>%  
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) 
```

### Creating scatterplot of boy names in 2016


```{r}
wboy_2016 = filter(bn_df, gender == "MALE" & year == 2016 & ethnicity == "WHITE_NH")

wboy_2016 %>% 
  ggplot() +
  geom_point(aes(x = rank, y = count), size  = 0.5) +
  geom_text(aes(rank, count, label = childs_first_name, angle = 5, hjust = -.1)) +
  theme_bw() +
  scale_y_continuous(breaks = seq(0, 300, 50)) +
  scale_color_viridis_c() +
  labs(title = "Ranked Non-Hispanic White Boy Names of 2016",
       x = "Rank of Name",
       y = "Number of Boys Born With Name")


#childs_first_name

```

